A partisan affair: Mapping patronage in municipal bureaucracies of Brazil

### 2021-01-15 paper/draft.tex
add municipal employment description

There is an ongoing debate on the strength of partisan ties in Brazil. On the one hand, scholars have noted that partisanship in Brazil is weak, meaning that politicians and voters do not have strong party loyalty and often ``switch'' to other parties \citet{desposato2006parties,ames2002deadlock}. On the other, some scholars have noted that party ties have grown in strength over time, in particular for leftist programmatic parties such as the PT \citep{samuels2014power,samuels2006sources}. Part of this debate owes to disagreements on how to measure party strength, whether it be testing voters' prior knowledge of party's ideological positions or if instead, it should be measured by testing whether voters issue ballots for individual candidates or their party labels -- with each one of these measures painting an opposite picture of the relative strength of party ties.

@@

\section{Descriptive Statistics}
\label{sec:descriptive}

>>>>

There is an ongoing debate on the strength of partisan ties in Brazil. On the one hand, scholars have noted that partisanship in Brazil is weak, meaning that politicians and voters do not have strong party loyalty and often ``switch'' to other parties \citep{desposato2006parties,ames2002deadlock}. On the other, some scholars have noted that party ties have grown in strength over time, in particular for leftist programmatic parties such as the PT \citep{samuels2014power,samuels2006sources}. Part of this debate owes to disagreements on how to measure party strength, whether it be testing voters' prior knowledge of party's ideological positions or if instead, it should be measured by testing whether voters issue ballots for individual candidates or their party labels -- with each one of these measures painting an opposite picture of the relative strength of party ties.

@@

Municipal employment in Brazil is under local jurisdiction, with personnel appointments under the exclusive authority of the executive branch. Salaries, contract modalities and terminations are also under municipal jurisdiction. There are multiple forms of contract available, a permanent contract (\emph{estatut\'{a}rio}), a regular contract (\emph{CLT}) and temporary hires. Personnel expenditures cannot exceed a ceiling of 60 percent of the local budget, but as long as the ceiling is respected, other levels of government cannot interfere with local personnel decisions.

Municipal discretion leaves ample room for patronage, whether it be to support mayoral coalitions as outlined in the first paper of this dissertation or to reward contributors and party loyalists \citep{colonnelli2018patronage,brollo2017victor}. These politically motivated appointments are particularly prominent in high-level positions, the so called \emph{cargos de confian\,{c}a}, local ministerial positions that are both high in compensation and provide access to decisions over key public services such as transportation and health.

\section{Descriptive statistics}
\label{sec:descriptive}
